Drawing Math

### 2022-05-13 paper/Paper.tex
clean up

\pagebreak
\section*{Abstract}
Given an object in cartesian space, we ask whether a repeating sequence 
of spherical updates to the object's position cause it to travel on a closed path.
A spherical update to a position in $D$-dimensional space can be written as
$\left(r, \theta_1, \theta_2, ..., \theta_{D-1}\right)$.
In this paper, we consider $D-1$ non-terminating
rational numbers where for time-step $i$, the spherical updates' difference function
is given by $r' \leftarrow r$, 
$\theta_1' \leftarrow \theta_1 + \mathrm{digit}(q_1, i)$, ..., 
$\theta_{D-1}' \leftarrow \theta_{D-1} + \mathrm{digit}(q_{D-1}, i)$.
We then proceed to derive a formula for finding if the object takes a closed, repeating path.
Moreover, we explore interesting properties of this problem and
relate it to discrete log, roots of a multinomial, and center of mass.

\section{Introduction}
A glum Pittsburgh day inspired the authors to take a random walk down YouTube's recommended
when they encountered the Numberphile video on ``Plotting Pi" \cite{NumberphileVid}.
In the video, Henderson and Brady introduce the idea of taking a Python Turtle and
deriving a series of updates to its position based off of various decimal sequences, 
some rational, some irrational. More specifically, they place an object in 2D, cartesian
space starting at $(0, 0)$ and ``facing" to the right. They then take a generating number, like $\pi$ or $35/99$, 
which gives a decimal sequence ($3.14...$ and $0.\bar{35}$ respectively). 
Then, at time step $i$, $i \geq 1$, they rotate the object by the $i$th digit divided by the base of
the decimal sequence. Then, the turtle moves a constant distance in the direction which it faces.
For example, if the second digit is 4, the turtle is rotated by $\frac{4}{10}$ths of a circle
counterclockwise at time step 2 and then moves 10 units in its new direction.
\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{fig/turtle2dtimestep.png}
  \caption{The turtle moving according to decimal sequence $0.223$ in base $10$.}
\end{figure}

Whenever the turtle moves, the program draws a red line along the turtle's path. Thus,
different lines are drawn.
The authors noticed that using rational numbers to generate the sequence often drew geometrically
aesthetic, closed shapes.

\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{fig/Turtle1D13Div113.png}
  \caption{The closed shape generated from $\frac{13}{113} = 0.0977443609...$ in base $10$.}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{fig/Turtle1DNotClosed.png}
  \caption{A non-closing shape generated from $\frac{2134}{9999} = 0.\overline{2134}$ in base $10$.}
\end{figure}

Naturally, the authors were curious whether a Turtle in ``3D" would also draw 
aesthetic shapes. In other words, what if the turtle's orientation was described
by 2 angles, pitch and yaw. The pitch and yaw would then by independently updated
by 2 decimal sequences generated from 2 rational numbers.
% TODO: figure
\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{fig/turtle.png}
  \caption{The above is a placeholder, I would like to get Shiva to help me design a 3D graphic here}
\end{figure}
After writing the program, we noticed that the turtle would often fail
to close. In other words, the turtle would go off in one direction forever.
% TODO: picture of failing to close.
But, the turtle \textit{sometimes} closed.

@@

% \section{Background}
% - Digital math
% - Euler's formula
% - I payed attention in some of my lectures

\pagebreak
\section{Definitions and questions}
\subsection{Definitions}
Say we had a turtle, Jimbo, living in $D$ dimensional Euclidean
space and in discrete time. At time step $i$, where $i \in \Z$ and $i > 0$,
the turtle has position $p_i \in \R^D$.
Then, lets define $\Delta p_{i+1} = p_{i+1} - p_i$; in other words, $\Delta p_{i+1}$ is the change in position from time $i$ to $i + 1$.

Now say that the turtle's movement is determined by $k$ seed parameter drawn from
the same set. Then, for some state space $\mathcal{S}$,
define $s_i^j \in \mathcal{S}$ to be some arbitrary
state associated with timestamp $i$ for the $j$th seed parameter where $j \in [k]$.
Also, define $\pmb{s}_i = (s_i^1, s_i^2, ..., s_i^k)$.
Next we will define a set of functions $SU^j: \mathcal{S} \rightarrow \mathcal{S}$
(for $S$tate $U$pdater) such that $s_{i + 1}^j = SU^j(s_i^j, i)$. Note that for
$j, a \in [m]$ where $j \neq a$, $s_{i+1}^j$ is determined solely by $s_i^j$ and $i$ and not
$s_i^a$.

@@

Finally, let us define
$$
  \Delta P_{a, b} = \sum_{i = a}^{b} \Delta p_i.
$$
In other words, $\Delta P_{a, b}$ is the change in position from timestep $a$ to $b$.

\subsection{The problem}
Say we are given, $Comb$, $SU^j$, $p_0$, and $s^j_0$ for all $j \in [k]$.
Informally, the question is whether the turtle draws a ``closed" shape or not.\\
More formally, is there some period $T$ such that 
$$
  p_{i + \ell T} = p_{i}
$$
for all $i, \ell \in \N$.
Then, note that if there exists a period $T$ such that 
$\Delta P_{i, i + \ell T} = 0$ for all $i, \ell \in N$, 
$p_{i + \ell T} = p_{i}$ and the turtle forms a closed shape.

\subsection{Specifying the task ahead of us}
For our case, we consider $Comb, SU_i^j$ to be uniquely determined by the current input.
So, we can simplify the overall question.
If, $\pmb{s}_i = \pmb{s}_{i + \ell T}$ for some $T \in \N$ and all $i \in \N$, then 
$\Delta p_{i} = \Delta p_{i + \ell T}$. So then,
$\Delta P_{i, i + \ell T} = \Delta P_{i, i + \ell' T}$ for all $\ell, \ell' \in \N$.
Thus, $T$ is a period of the change in position. We can thus
break down our problem into two parts:
\begin{enumerate}
  \item Finding the period, $T$, of the state $\pmb{s}$.
  \item Checking whether $\Delta P_{i, i + T} = 0$.
\end{enumerate}

\subsection{Some more restrictions on our problem}
We further restrict the problem by only considering $\mathcal{S} = \N^4$ where for $(n, d, b, \theta) \in \mathcal{S}$,
$n$ is the numerator of a rational in fraction form, $d$ is the denominator, 
$b$ is the base (i.e.\ base 10, base 12, etc.),
and $\frac{2\theta}{b \pi}$ is an ``angle" associated with the state.

@@

Moreover, we consider the case where
$$
  Comb((., ., ., ., \theta^1), (., ., ., ., \theta^2), ..., (., ., ., ., \theta^k)) =
    \left(\combSingleTerm{1}, ..., \combSingleTerm{D}\right)
$$
where $\mathrm{incl}_d^j \in \set{0, 1}$ for $d \in [D]$ indicates whether to include a given
$x \in R$ determined by rational parameter $j$ for position update in the $d$th dimension.

Finally, for simplicity's sake, assume that $\theta = 0$ for all $(n, b, d, \theta) \in \pmb{s}_0$,
$n < d$, and $\frac{n}{d}$'s decimal expansion is periodic after some $N \geq 0$ decimal places
and does not terminate in base $b$.

@@

\subsubsection{Some intuition}
While the restrictions may seem arbitrary, they aptly match our original problem statement.
The original problem statement derives a spherical change in position based off of a rational number's
digit at a particular timestep. The polar change in position also has a fixed radius. Translating
from a polar to cartesian update then only requires products of $\sin$s and $\cos$s. See \cite{NDimSphericalCoord}
for more details.

@@

\subsection{Finding period $T$}
We now show how to find a period $T$ of state $\pmb{s}$.

\subsubsection{Finding the period of a fraction}
For some $(n, b, d, \theta) \in \pmb{s}_0$,
we have that the period of the decimal expansion of $\frac{n}{d}$
can be determined by finding the smallest $\rationalPeriod$ such that
\begin{equation}
\label{eq:nd-period}
b ^ {\rationalPeriod} \equiv 1 \mod d
\end{equation}
by \cite{MathOverflowFracPeriod}.
More generally though, any nontrivial $\rationalPeriod$ satisfying equation \ref{eq:nd-period}
will be a period of $\frac{n}{d}$. In other words, the decimal sequence specified 
by $\frac{n}{d}$ will repeat after every $\rationalPeriod$ steps for all digits 
$n > n_0$ for some fixed $n_0$.

@@

\subsubsection{Digital sum}
Next, we introduce the idea digital sums.
For some number $N \in \N$, $N$ can be represented in base $b$ via
\begin{equation}
  N = \sum_{i=0}^{m} d_i b^i
\end{equation}
where $m = \lceil \log_b N \rceil$ and, $\forall i \in [m]$, $d_i \in \Z_b$.
Then, we define function $\mathrm{digSum}: \N \rightarrow \Z_b$
to give the digital sum such that
\begin{equation}
  \mathrm{digSum}(N) = \sum_{i=0}^{m} d_i.
\end{equation}

Moreover, define $\digSumPeriodic \in \Z_b$
such that
% TODO: define T within the func from another func
% TODO: bettr defn
\begin{equation}
  \digSumPeriodic = \sum_{i = i_0}^{i_0 + \rationalPeriodTotal} \mathrm{digit}(n, d, b, i).
\end{equation}
In other words, $\rationalPeriodTotal$ is the digital sum over one period.

\begin{remark}[Complexity]
\label{remark:digsum-complexity}
  For $d > 2$, prime, and coprime to $b$, we can find $\digSumPeriodic$
  in polytime by multiplying $(b - 1) \cdot \frac{d-1}{2} \mod b$ \cite{OnDecSeq}. The authors are unsure
  as to the complexity of finding $\digSumPeriodic$ otherwise.
\end{remark}

\subsubsection{Finding a period of $\rationalAngleNotime$} % TODO: redifine things in terms of theta...
For $(n^j, b^j, d^j, \theta_i^j) = s_i^j$,
recall that $\rationalAngle{i + 1} = \rationalAngle{i} + \mathrm{digit}(n, b, d, i) \mod b$.
So, after period $\rationalPeriodTotal$,
\begin{align*}
  \theta_{i + \rationalPeriodTotal} &= \left(\theta_{i} + \sum_{\ell = i}^{\rationalPeriodTotal + i} \mathrm{digit}(n, b, d, \ell)\right) \mod b\\
  &= \left(\theta_{i} +  \digSumPeriodic\right) \mod b.
\end{align*}
So, after $p$ periods of length $\rationalPeriodTotal$ where $p  \cdot \digSumPeriodic \equiv 0 \mod b$,
$$
\theta_{i + p\rationalPeriodTotal} \equiv \theta_i + 0 \equiv \theta_i.
$$
For simplicity, lets define
$$
  T^j = p\rationalPeriodTotal
$$
where $T^j$ is a period of the state for rational parameter $j$.
% TODO: remove $i$ from the state and somehow have it be implicit....

% Oh man... we are going to have to define an intermediate total period...
% One for all T'...
% and then have a sum for T
\subsubsection{Finding the period of $\pmb{s}$}
We can first see that for $\stateNotime{j} \in \pmb{s}$, $\stateNotime{j}$
has period of $\statePeriod$. So, $\pmb{s}$ must have a period, $\totalPeriod$, of
$$
\lcm{j \in [k]}\; T^j.
$$
I.e.\ $\pmb{s}_i = \pmb{s}_{i + T}$ for all $i \in \N$.

% TODO: cite https://math.stackexchange.com/questions/377683/length-of-period-of-decimal-expansion-of-a-fraction
% TODO: formula in poly time
% TODO: hmmm... specify i > n_0 (to get rid of initial stuff...)

% https://en.wikipedia.org/wiki/Digital_root#:~:text=The%20digital%20root%20(also%20repeated,single%2Ddigit%20number%20is%20reached.

\subsection{Finding the change in position over a period}
So now that we know the period of $\totalStateNotime$, we can ask if
$\Delta P_{i, i + T} = 0$. % TODO: reference

@@

\section{Interesting Properties and an attempt at certificates}
We will now proceed to go over some interesting properties of the closure question
which may give rise to an algorithm in NP, coNP, or even BQP. These properties were discovered
in the author's pursuit of simplifying the question. Moreover, these properties may guide
some intuition as to the probability of closure for random rational seeds, a fixed $k$, 
and fixed bases $b$.

@@

Then, let
$$
\mathcal{M} = \{f : f(\pmb{x}) = \pm x_1 \pm x_2 ... \pm x_{|A_d|}\}.
$$
In other words, $\mathcal{M}$ is the set of all multinomials with $|A_d|$ variables with
degree 1 and coefficients $\pm1$.
Then,
if \begin{equation}
\label{eq:satisfy-multi}
f\left(\digSumPeriodicNoJ\right) \neq 0
\end{equation}
for all $f \in \mathcal{M}$ and $d \in [D]$, the turtle will always draw a closed shape. See 
Appendix \ref{appendix:prop1} for the proof.

Satisfying \eqref{eq:satisfy-multi} is true for all $f$ is equivalent to
$$
  \prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) \neq 0.
$$

@@

Then, let
$$
  \mathcal{B} = \left\{\beta: f_\beta(\sigma) = 0\right\}
$$
and $\overline{\mathcal{B}} = \mathcal{M} \setminus \mathcal{B}$. Then, if for all
$\beta \in \mathcal{B}$,
$$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
  \exp\left(
      I 
      \commonToAngle 
      \sum_{\ell = 0}^{q}
      \sum_{j \in A_d}
          \left(-1\right) ^ {\beta_{j}}
          \numbToCommon
          \mathrm{digit}(n, b, d, \ell)
  \right) = 0
$$
the turtle will draw a closed shape. While this may seem arbitrary, there is a unique
and interesting geometric interpretation. 
Let a rational decimal sequence $a_1, a_2, ..., a_{\rationalPeriodTotal}$ be equal to
$
\forall \ell \in [\rationalPeriodTotal]
$,
$$
a_\ell = \sum_{j \in A_d}
\left(-1\right) ^ {\beta_{j}}
\numbToCommon
\mathrm{digit}(n, b, d, \ell).
$$
In other words, we are creating a ``common" rational sequence by summing and subtracting
our rational parameters in the least common multiple base $\commonBase$. Then,
if $$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
\exp\left(
    I 
    \commonToAngle 
    \sum_{\ell = 0}^{q}
      a_\ell
  \right) = 0
$$
the turtle closes.
If $r_i = \commonToAngle \sum_{\ell = 0}^{q} a_\ell$ for 
$i \in \set{0, ..., \rationalPeriodTotal - 1}$, then $\exp(I \cdot r_i)$ can be thought of as some point at angle $r_i$ around the 
complex unit circle. Also, note that $r_{i + 1} = r_i + \commonToAngle a_{i+1}$. In otherwords,
$r_i$ denotes some position around the complex unit circle where the subsequent updates
to position are given the $a_i$s, which are a sum and subtraction of rational parameters. Then, if the center of
mass of these points around the unit circle is $0$, the turtle draws a closed shape!
The authors are unsure as to the underlying intuition for why this is true.

% TODO: picture!!!!

\section{Open Questions and Future Work}
A whole host of questions naturally arise from this problem. We will proceed to list
the ones which immediately stand out. For a complete list of open problems, which have
been solved, and the associated (financial) reward to each question, see our website
\begin{enumerate}
  \item \textbf{Periodic Digit Sum in NP?}: As far as the authors are aware, given 
  any rational number, $\frac{n}{d}$ in a fixed base $b$, is finding the sum of the digits in
  one period modulo $b$ in $NP$? I.e. could one give a poly-size
  certificate to a poly-time verifier which verifies if a digital sum equals $\sigma \in \Z_b$?
  The size of the input is the number of bits required to describe $\frac{n}{d}$. The authors 
  feel that intuitively the answer should be yes. A certificate could potentially be related
  to the factors of the denominator and the period length. I.e.\ we know how to find
  the periodic digital sum of a prime in polynomial time (see \ref{remark:digsum-complexity}). So,
  could we somehow use this in combination with a prime factorization?

  \item \textbf{Closure in NP or CoNP?}: For fixed bases of size $b$, is closure in NP of CoNP?
  I.e.\ is there is a poly-sized certificate and poly-time verifier which can prove or disprove
  that $D-1$ rationals draw a closed shape? The author's are uncertain here but conjecture
  may be in CoNP. Either a gadget based reduction or a direct algorithm to generate a valid
  certificate suffice to answer this question.

@@

  \item \textbf{Decomposition}: This one is totally out of left field. From our analysis,
  it seems as if using rational numbers to draw out shapes has a lot of underlying complexity
  but also a lot of structure. The authors were wondering if given any closed shape $\ell$ which
  can be described with $n$ bits, could poly-n bits be used in combination with the rational
  sequence drawing algorithm described throughout this paper as description of $\ell$.
  In other words, could some sort of ``Fourier Transform" be done between an explicitly described
  closed shape and a sequence of rationals? We are also interested whether $1 - \epsilon$
  approximation exists 
  (where the approximation ``goodness" could be measured as a function of
  average distance from the original line to the line generated via the sequence of rationals).

  \item \textbf{Closure in BQP?}: Given that closure is intimately tied with period finding,
  the authors wonder if this problem is in BQP (Bounded-error Quantum Polynomial Time). An algorithm
  or existence result suffices.

>>>>

\pagebreak
\section*{Abstract}
Given an object in $D$ dimensional cartesian space, we ask whether a repeating sequence 
of deltas to the object's position cause it to travel on a closed path.
We consider deltas specified in a spherical basis which can be written as
$\left(r, \theta_1, \theta_2, ..., \theta_{D-1}\right)$.
In this paper, we consider $D-1$ non-terminating
rational numbers where for time-step $i$, the deltas are given by $r' \leftarrow r$, 
$\theta_1' \leftarrow \theta_1 + \mathrm{digit}(q_1, i)$, ..., 
$\theta_{D-1}' \leftarrow \theta_{D-1} + \mathrm{digit}(q_{D-1}, i)$.
We then proceed to derive a formula for finding if the object takes a closed, repeating path.
Moreover, we explore interesting properties of this problem and
relate it to discrete log, roots of a multinomial, and center of mass.

\section{Introduction}
A glum Pittsburgh day inspired the authors to take a random walk down YouTube's recommended section
when they encountered the Numberphile video on ``Plotting Pi" \cite{NumberphileVid}.
In the video, Henderson and Brady introduce the idea of taking a Python Turtle and
deriving a series of updates to its position based off of various decimal sequences, 
some rational, some irrational. More specifically, they place an object in 2D, cartesian
space starting at $(0, 0)$ and ``facing" to the right. They then take a generating number, like $\pi$ or $35/99$, 
which gives a decimal sequence ($3.14...$ and $0.\overline{35}$ respectively). 
Then, at time step $i$, $i \geq 1$, they rotate the object by the $i$th digit divided by the base of
the decimal sequence. Then, the turtle moves a constant distance in the direction which it faces.
For example, if the second digit is 4, the turtle is rotated by $\frac{4}{10}$ths of a circle
counterclockwise at time step 2 and then moves 10 units in its new direction.
\begin{figure}[h!]
  \centering
  \includegraphics[width=8cm]{fig/turtle2dtimestep.png}
  \caption{The turtle moving according to decimal sequence $0.223$ in base $10$.}
\end{figure}

Whenever the turtle moves, the program draws a red line along the turtle's path.
The authors noticed that using rational numbers to generate the sequence often drew geometrically
aesthetic, closed shapes.

\begin{figure}[h!]
  \centering
  \includegraphics[width=8cm]{fig/Turtle1D13Div113.png}
  \caption{A closed 2D shape generated from $\frac{13}{113} = 0.0977443609...$ in base $10$.}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=8cm]{fig/Turtle1DNotClosed.png}
  \caption{A non-closing 2D shape generated from $\frac{2134}{9999} = 0.\overline{2134}$ in base $10$.}
\end{figure}

Naturally, the authors were curious whether a Turtle in ``3D" would also draw 
interesting paths. In other words, what if the turtle's orientation was described
by 2 angles, pitch and yaw. The pitch and yaw would then by independently updated
by 2 decimal sequences generated from 2 rational numbers.
% TODO: figure
\begin{figure}[h!]
  \centering
  \includegraphics[width=8cm]{fig/turtle.png}
  \caption{The above is a placeholder, I would like to get Shiva to help me design a 3D graphic here similar to Figure 1 but in 3D}
\end{figure}
After writing the program, we noticed that the turtle would often fail
to close. In other words, the turtle would go off in one direction forever.
% TODO: picture of failing to close in 3d space
But, the turtle \textit{sometimes} closed.
\begin{figure}[h!]
  \centering
  \includegraphics[width=8cm]{fig/failclose3d.png}
  \caption{A non closing 3D shape generated from $\frac{102}{999} = 0.\overline{102}$ and $\frac{223}{999} = 0.\overline{223}$ in base 10.}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=8cm]{fig/tempclose.jpeg}
  \caption{A closing 3D shape generate from (?????... this is a placeholder).}
\end{figure}

@@

% \section{Background}
% - Digital math
% - Euler's formula
% - I payed attention in some of my lectures
\pagebreak

\section{Definitions and questions}
\subsection{Definitions}
Say we had a turtle, Jimbo, living in $D$ dimensional Euclidean
space and in discrete time. At time step $i$, where $i \in \Z$ and $i > 0$,
the turtle has position $p_i \in \R^D$.
Then, lets define $\Delta p_{i+1} = p_{i+1} - p_i$; in other words, $\Delta p_{i+1}$ is the change in position from time $i$ to $i + 1$.

Now say that the turtle's movement is determined by $k$ parameter drawn from
the same set, $\mathcal{P}$. Then, for some state space $\mathcal{S}$,
define $s_i^j \in \mathcal{S}$ to be some arbitrary
state associated with time-step $i$ for the $j$th seed parameter where $j \in [k]$.
Also, define $\pmb{s}_i = (s_i^1, s_i^2, ..., s_i^k)$.
Next we will define a set of functions $SU^j: \mathcal{S} \rightarrow \mathcal{S}$
(for $S$tate $U$pdater) such that $s_{i + 1}^j = SU^j(s_i^j, i)$. Note that for
$j, a \in [m]$ where $j \neq a$, $s_{i+1}^j$ is determined solely by $s_i^j$ and $i$ and not
$s_i^a$.

@@

Finally, let us define
$$
  \Delta P_{a, b} = \sum_{i = a}^{b} \Delta p_i.
$$
In other words, $\Delta P_{a, b}$ is the change in position from time-step $a$ to $b$.

\subsection{The problem}
Say we are given, $Comb$, $SU^j$, $p_0$, and $s^j_0$ for all $j \in [k]$.
Informally, the question is whether the turtle draws a ``closed" shape or not.\\
More formally, is there some period $T$ such that 
$$
  p_{i + \ell T} = p_{i}
$$
for all $i, \ell \in \N$.
Then, note that iff there exists a period $T$ such that 
$\Delta P_{i, i + \ell T} = 0$ for all $i, \ell \in N$, 
$p_{i + \ell T} = p_{i}$ and the turtle forms a closed shape.

\subsection{Specifying the task ahead of us}
For our case, we consider $Comb, SU_i^j$ to be uniquely determined by the current state.
So, we can simplify the overall question.
If, $\pmb{s}_i = \pmb{s}_{i + \ell T}$ for some $T \in \N$ and all $i \in \N$, then 
$\Delta p_{i} = \Delta p_{i + \ell T}$. So then,
$\Delta P_{i, i + \ell T} = \Delta P_{i, i + \ell' T}$ for all $\ell, \ell' \in \N$.
Thus, $T$ is a period of the change in position. We can thus
break down our problem into two parts:
\begin{enumerate}
  \item Finding the period, $T$, of the state $\pmb{s}$.
  \item Checking whether $\Delta P_{i, i + T} = 0$.
\end{enumerate}

\subsection{Some more restrictions on our problem}
We further restrict the problem by only considering $\mathcal{S} = \N^4$ where for $(n, d, b, \theta) \in \mathcal{S}$,
$n$ is the numerator of a rational in fraction form, $d$ is the denominator, 
$b$ is the base,
and $\frac{2 \pi \theta}{b}$ is an ``angle" associated with the state.

@@

Moreover, we consider the case where
$$
  Comb((n^1, d^1, b^1, \theta^1), (n^2, b^2, b^2, \theta^2), ..., (n^k, d^k, b^k, \theta^k)) =
    \left(\combSingleTerm{1}, ..., \combSingleTerm{D}\right)
$$
where $\mathrm{incl}_d^j \in \set{0, 1}$ for $d \in [D]$ indicates whether to ``include" a
rational parameter $j$ for position update in the $d$th dimension.

Finally, for simplicity's sake, assume that $\theta = 0$ for all $(n, b, d, \theta) \in \pmb{s}_0$,
$n < d$, $\frac{n}{d}$'s decimal expansion is periodic after some $N \geq 0$ decimal places,
and the decimal expansion does not terminate in base $b$.

@@

\subsubsection{Some intuition}
While the restrictions may seem arbitrary, they aptly match our original problem statement.
The original problem statement derives a spherical change in position based off of a rational number's
digit at a particular time-step. The spherical change in position also has a fixed radius. Translating
from a spherical to cartesian update then only requires products of $\sin$'s and $\cos$'s. See \cite{NDimSphericalCoord}
for more details.

@@

\subsection{Finding the state's period}
We now show how to find a period $T$ of state $\pmb{s}$.

\subsubsection{Finding the period of a decimal sequence}
For some $(n, b, d, \theta) \in \pmb{s}_0$,
we have that the period of the decimal expansion of $\frac{n}{d}$
can be determined by finding the smallest $\rationalPeriod$ such that
\begin{equation}
\label{eq:nd-period}
b ^ {\rationalPeriod} \equiv 1 \mod d
\end{equation}
by \cite{MathOverflowFracPeriod}.
More generally though, any nontrivial $\rationalPeriod$ satisfying equation \ref{eq:nd-period}
will be a period of $\frac{n}{d}$. In other words, the decimal sequence specified 
by $\frac{n}{d}$ will repeat after every $\rationalPeriod$ steps for all digits 
$n > n_0$ for some fixed $n_0$.

@@

\subsubsection{Digital sum}
Next, we introduce the idea of digital sums.
For some number $N \in \N$, $N$ can be represented in base $b$ via
\begin{equation}
  N = \sum_{i=0}^{m} d_i b^i
\end{equation}
where $m = \lceil \log_b N \rceil$ and, $\forall i \in [m]$, $d_i \in \Z_b$.
Then, we define function $\mathrm{digSum}: \N \rightarrow \Z_b$
to give the digital sum such that
\begin{equation}
  \mathrm{digSum}(N) = \sum_{i=0}^{m} d_i.
\end{equation}

Moreover, define $\digSumPeriodic \in \Z_b$
such that
% TODO: define T within the func from another func
% TODO: bettr defn
\begin{equation}
  \digSumPeriodic = \sum_{i = i_0}^{i_0 + \rationalPeriodTotal} \mathrm{digit}(n, d, b, i).
\end{equation}
In other words, $\digSumPeriodic$ is the digital sum over 
a common decimal sequence period for all rational parameters.

\begin{remark}[Complexity]
\label{remark:digsum-complexity}
  For $d > 2$, prime, and coprime to $b$, we can find $\digSumPeriodic$
  in polytime by multiplying $\frac{\commonBase}{b}(b - 1) \cdot \frac{d-1}{2} \mod b$ \cite{OnDecSeq}. The authors are unsure
  as to the complexity of finding $\digSumPeriodic$ otherwise.
\end{remark}

\subsubsection{Finding a period of a parameter's state} % TODO: redifine things in terms of theta...
For $(n^j, b^j, d^j, \theta_i^j) = s_i^j$,
recall that $\rationalAngle{i + 1} = \rationalAngle{i} + \mathrm{digit}(n, b, d, i) \mod b$.
So, after period $\rationalPeriodTotal$,
\begin{align*}
  \theta_{i + \rationalPeriodTotal} &= \left(\theta_{i} + \sum_{\ell = i}^{\rationalPeriodTotal + i} \mathrm{digit}(n, b, d, \ell)\right) \mod b\\
  &= \left(\theta_{i} +  \digSumPeriodic\right) \mod b.
\end{align*}
So, after $p$ periods of length $\rationalPeriodTotal$ where $p  \cdot \digSumPeriodic \equiv 0 \mod b$,
$$
\theta_{i + p\rationalPeriodTotal} \equiv \theta_i + 0 \equiv \theta_i.
$$
Now, lets define
$$
  T^j = p\rationalPeriodTotal
$$
where $T^j$ is a period of the state for rational parameter $j$.

% Oh man... we are going to have to define an intermediate total period...
% One for all T'...
% and then have a sum for T
\subsubsection{Finding the period of the overall state}
We can first see that for $\stateNotime{j} \in \pmb{s}$, $\stateNotime{j}$
has period of $\statePeriod$. So, $\pmb{s}$ must have a period $\totalPeriod$ equaling
$$
\lcm{j \in [k]}\; T^j.
$$
I.e.\ $\pmb{s}_i = \pmb{s}_{i + T}$ for all $i \in \N$.

\subsection{Finding the change in position over a period}
So now that we know the period of $\totalStateNotime$, we can ask if
$\Delta P_{i, i + T} = 0$.

@@

\section{Interesting Properties}
We will now proceed to go over some interesting properties of the closure question
which may give rise to an algorithm in NP, coNP, or even BQP. These properties were discovered
in the author's pursuit of simplifying the question. Moreover, these properties may guide
some intuition as to the probability of closure for random rational seeds, a fixed $k$, 
and fixed bases $b$.

@@

Then, let
$$
\mathcal{M} = \{f : f(\pmb{x}) = \pm x_1 \pm x_2 ... \pm x_{|A_d|}\}.
$$
In other words, $\mathcal{M}$ is the set of all multinomials with $|A_d|$ variables with
degree 1 and coefficients $\pm1$.
Then,
if \begin{equation}
\label{eq:satisfy-multi}
f\left(\digSumPeriodicNoJ\right) \neq 0
\end{equation}
for all $f \in \mathcal{M}$ and all dimensions, the turtle will always draw a closed shape. See 
Appendix \ref{appendix:prop1} for the proof.

Then, satisfying \eqref{eq:satisfy-multi} for all $f$ in a dimension is equivalent to
$$
  \prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) \neq 0.
$$

@@

Then, let
$$
  \mathcal{B} = \left\{\beta: f_\beta(\sigma) = 0\right\}.
$$
Then, if for all
$\beta \in \mathcal{B}$,
$$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
  \exp\left(
      I 
      \commonToAngle 
      \sum_{\ell = 0}^{q}
      \sum_{j \in A_d}
          \left(-1\right) ^ {\beta_{j}}
          \numbToCommon
          \mathrm{digit}(n, b, d, \ell)
  \right) = 0
$$
the turtle will draw a closed shape. While this may seem arbitrary, there is a unique
and interesting geometric interpretation. 
Let a rational decimal sequence $a_1, a_2, ..., a_{\rationalPeriodTotal}$ be equal to
$
\forall \ell \in [\rationalPeriodTotal]
$,
$$
a_\ell = \sum_{j \in A_d}
\left(-1\right) ^ {\beta_{j}}
\numbToCommon
\mathrm{digit}(n, b, d, \ell).
$$
In other words, we are creating a ``common" rational sequence by summing and subtracting
our rational parameters in base $\commonBase$. Then,
if $$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
\exp\left(
    I 
    \commonToAngle 
    \sum_{\ell = 0}^{q}
      a_\ell
  \right) = 0
$$
the turtle closes.
If $r_i = \commonToAngle \sum_{\ell = 0}^{q} a_\ell$ for 
$i \in \set{0, ..., \rationalPeriodTotal - 1}$, then $\exp(I \cdot r_i)$ can be thought of as some point at angle $r_i$ around the 
complex unit circle. Also, note that $r_{i + 1} = r_i + \commonToAngle a_{i+1}$. In otherwords,
$r_i$ denotes some position around the complex unit circle where the subsequent updates
to position are given the $a_i$s, which are a sum and subtraction of rational parameters. Then, if the center of
mass of these points around the unit circle is $0$, the turtle draws a closed shape!
The authors are unsure as to the underlying intuition for why this is true.

For example, say that rational parameters $\frac{22,022}{99,999} = 0.\overline{22022}$ and
$\frac{200}{99,999} = 0.\overline{00200}$ determine the movement in the $x$, $y$, and $z$ axis for $3$ dimensions.
Then, $\sigma^1 = 8$ and $\sigma^2 = 2$. Also note that we are in base 10 so $b^1 = b^2 = \commonBase = 10$.
Also, $f_{00}(8, 2) = 8 + 2 = 10 \equiv 0 \mod 10$. But then, we have that $a = 0.\overline{22222}$.
So then, the ``movement" around the unit circle ends up being evenly spaced out and the center of mass
is 0. Thus, we know that the turtle will close.

\begin{figure}[h!]
  \centering
  \includegraphics[width=8cm]{fig/evenspacing.png}
  \caption{If $a = 0.\overline{22222}$, then $r_1, ..., r_5$ are spaced around the unit circle as shown in the figure. We can thus
  see that the center of mass is 0.}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=8cm]{fig/3DTurtleCenterOfMass.png}
  \caption{The turtle moving according to rationals $\frac{22,022}{99,999}$ and $\frac{200}{99,999}$.}
\end{figure}

\pagebreak
\section{Open Questions and Future Work}
A whole host of questions naturally arise from this problem. We will proceed to list
the ones which immediately stand out. For a complete list of open problems, which have
been solved, and the associated (financial) reward to each question, see our website
\begin{enumerate}
  \item \textbf{Periodic Digit Sum in NP?}: As far as the authors are aware, given 
  any rational number, $\frac{n}{d}$ in a fixed base $b$, is finding the sum of the digits in
  one period modulo $b$ in $NP$? I.e. could one give a poly-size
  certificate to a poly-time verifier which verifies if a digital sum equals $\sigma \in \Z_b$?
  The size of the input is the number of bits required to describe $\frac{n}{d}$. The authors 
  feel that intuitively the answer should be yes. A certificate could potentially be related
  to the prime factors of the denominator and the period length. I.e.\ we know how to find
  the periodic digital sum of a prime in polynomial time (see \ref{remark:digsum-complexity}). So,
  could we somehow use this in combination with a prime factorization?

  \item \textbf{Closure in NP or CoNP?}: For fixed bases of size $b$, is closure in NP of CoNP?
  I.e.\ is there is a poly-sized certificate and poly-time verifier which can prove or disprove
  that $D-1$ rationals draw a closed shape? The author's are uncertain here but conjecture
  may be in CoNP.

@@

  \item \textbf{Decomposition}: This one is totally out of left field. From our analysis,
  it seems as if using rational numbers to draw out shapes has a lot of underlying complexity
  but also a lot of structure. The authors were wondering if given any closed shape $\ell$ which
  can be described with $n$ bits, could poly-n bits be used in combination with the rational
  sequence drawing algorithm described throughout this paper as description of $\ell$.
  In other words, could some sort of ``Fourier Transform" be done between an explicitly described
  closed shape and a sequence of rationals? We are also interested whether $1 - \epsilon$
  approximation exists 
  (where the approximation factor could be measured as a function of
  average distance from the original line to the line generated via the sequence of rationals).

  \item \textbf{Closure in BQP?}: Given that closure is intimately tied with period finding,
  the authors wonder if this problem is in BQP (Bounded-error Quantum Polynomial Time).
